Logsums

### 2019-07-15 03_logsums.Rmd
Revise logsum function
It's now more simple with a basic impedance and size term.

## Logsums
Here's the function we use to calculate the logsums
```{r}
#' Calculate destination choice logsums from a distance matrix
#'
#' @param d An $n\times p$ matrix with the distance from all tracts to
#'   all parks
#' @param sizes A p-length vector of park sizes
#' @param tweets A p-length vector of tweets at parks
#' @param betas A vector containing the size, distance, and tweet coefficients.
#'   If we submit two variables the tweets are ignored.
#'
#' @return An n-length vector containing the weighted log-sum based
#'   accessibility between a tract and all parks.
#' @details If we have n tracts and p parks, distances needs to be a
#'
calculate_park_logsums <- function(d, sizes, tweets = NULL,
                                   betas = c(-0.4,0.2,0.001)){

  # A is n x p
  a <- betas[1] * d

  # B is p x 1
  b <- betas[2] * sizes

  if(!is.null(tweets))
    b <- b + betas[3] * tweets

  # calculate observed utility by adding the weighted park-level attributes
  # to the columns of the matrix
  # V is n x p, with b added by-column to each element in a
  V <- sweep(a, 2, b, `+`)

  # log-sum of exponentiated utility, Output is n-length vector
  log(rowSums(exp(V)))

}
```

>>>>

## Logsums

Here's the function we use to calculate the logsums. 
```{r}
#' Calculate destination choice logsums from a distance matrix
#'
#' @param impedance An $n\times p$ matrix with the distance from all tracts to
#'   all parks
#' @param size_term A p-length vector of the park size terms
#'
#' @return An n-length vector containing the weighted log-sum based
#'   accessibility between a tract and all parks.
#' @details If we have n tracts and p parks, distances needs to be a
#'
calculate_park_logsums <- function(impedance, size_term){

  # calculate observed utility by adding the weighted park-level attributes
  # to the columns of the matrix
  # V is n x p, with b added by-column to each element in a
  V <- sweep(impedance, 2, size_term, `+`)

  # log-sum of exponentiated utility, Output is n-length vector
  log(rowSums(exp(V)))
}
```

```{r gravity}
logsum_vars <- tracts %>%
  transmute(
    geoid, 
    # Log-sum based utilities, without and with Twitter data
    access_ls = calculate_park_logsums(
      log(distances), open_spaces$acres, betas = c(-0.5, 0.05)),
    # Normalize the logsum onto the Z-scale
    access_ls_n = (access_ls - mean(access_ls) ) / sd(access_ls)
  ) %>%
  st_set_geometry(NULL)
```

```{r binary}
tracts_access <- tracts %>%
  left_join(logsum_vars, by = "geoid") %>%
  mutate(access_025 = ifelse(geoid %in% inside_025$geoid, TRUE, FALSE))

tracts_sp <- as(tracts_access, "Spatial")
```

```{r}
# this interactive map is much easier to create and explore
pal <- colorQuantile("GnBu", tracts_access$access_ls_n, n = 5)
bin <- colorFactor("Dark2", tracts_access$access_025)
leaflet(tracts_access) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(group = "access", color = ~ pal(access_ls_n),
              label = ~as.character(round(access_ls_n,2))) %>%
  # addPolygons(group = "tweets", color = ~ pal(tweets_ls),
  #             label = ~as.character(round(tweets_ls,2))) %>%
  addPolygons(group = "1/4 mile", color = ~ bin(access_025)) %>%
  addLayersControl(baseGroups = c("access", "tweets", "1/4 mile"))
```
